---
title: "Scraping Weather Data for Bordeaux and Antibes "
author: Antoine.P _ August 2017
output: html_notebook
---
We are going to use the package 'weatherData' in R to scrap one of Data souces  we need to keep building the Context Aware Reccomendation System for restaurant sector.

Firstly, we try to do some weather data analysis with using this package which can be founded at: "https://ram-n.github.io/weatherData/example_weatherYear.html"


( 2nd solution with PYTHON : Wunderground Data with Python Pandas & Seaborn Tutorial /Shane Lynn)

```{r}
library(weatherData)
city1 <- "BOD"
df1 <- getWeatherForYear(city1,2016)# this code can be used when we want to get a daily data set for one year
```

```{r}
head(df1,3)
```

```{r}
city2 <- "NCE"
df2 <- getWeatherForYear(city2, 2016)
```

```{r}
head(df2,3)
```

```{r}
a<-df1$Date
b<-df2$Date
manque_data<-b[!(b %in% a)]# to find out records in b but not in a
df22<-subset(df2,!(Date %in% c('2016-04-03','2016-04-04', '2016-05-12', '2016-10-29' ,'2016-11-19', '2016-11-20',  '2016-11-21')))

```

```{r}
getDailyDifferences <- function(df1, df22){
  Delta_Means <- df1$Mean_TemperatureC - df22$Mean_TemperatureC
  Delta_Max <- df1$Max_TemperatureC - df22$Max_TemperatureC
  Delta_Min <- df1$Min_TemperatureC - df22$Min_TemperatureC

  
  diff_df <- data.frame(Date=df1$Date, Delta_Means, Delta_Max, Delta_Min)
  return(diff_df)
}
```

```{r}
plotDifferences <- function (differences, city1, city2) {
library(reshape2)
  m.diff <- melt(differences, id.vars=c("Date"))
  p <- ggplot(m.diff, aes(x=Date, y=value)) + geom_point(aes(color=variable)) +  
    facet_grid(variable ~ .) +geom_hline(yintercept=0)
  p <- p + labs(title=paste0("Daily Temperature Differences: ", "Bordeaux", " minus ","Nice"))
  print(p)
}
```

```{r, echo=TRUE}
library(ggplot2)
differences<- getDailyDifferences(df1, df22)
plotDifferences(differences, city1, city2)
```
SCRAPING DATA FOR OUR PROJECT

First, get the station code of Bordeaux
```{r}
getStationCode("Bordeaux")
```
The code of Bordeaux is 'LFBD'.


Now, which columns relate to our project?


The code following will fetch the data by defaut, so giving us a data set containing 2 columns: TemperatureC and DateUTC

```{r}
bor_rawdat<-getWeatherForDate("LFBD",start_date = "2016-01-01",end_date = "2017-08-30",opt_detailed = TRUE)
```

However we are also interested in columns: 2,4,7,12,14. 

Please note that the column numbers are very different depending on whether or not opt_detailed is TRUE or FALSE.

When fetching the data, we have to set opt_custom_columns=TRUE and custom_columns=c(2,4,,7,12,14).

We are now ready to actually fetch the data.


```{r}
bor_rawdat<-getWeatherForDate("LFBD",start_date = "2016-01-01",end_date = "2017-08-30",opt_detailed = TRUE,opt_custom_columns=T, custom_columns=c(2,4,7,12,14))
```

```{r}
head(bor_rawdat)
```

```{r}
dim(bor_rawdat)
```

DOING THE SAME ACTION FOR ANTIBES


```{r}
getStationCode("Antibes")
```
IBONSON2

```{r}
showAvailableColumns("Antibes", start_date = "2016-01-01",end_date = "2017-08-30" )
```

```{r}
antibes_rawdat<-getWeatherForDate("Antibes",start_date = "2016-01-01", end_date = "2017-08-30",opt_detailed = TRUE,opt_custom_columns=TRUE,custom_columns = c(2,4,7,12,14))
```


```{r}
dim(antibes_rawdat)
```

```{r}
head(antibes_rawdat,3)
```

```{r}
tail(antibes_rawdat,3)
```

```{r}
getwd()
```
EXPORTING THESE 2 DATA SETS TO CSV FILES
```{r}
write.csv(bor_rawdat,"Bordeaux_weather_data.csv",row.names = FALSE)

```

```{r}
write.csv(antibes_rawdat,"Antibes_weather_data.csv", row.names = FALSE)
```

DATA VISUALIZATION AND EXPLORATION

```{r}
BORdf<-read.csv("C:/Users/Pham Antoine/Bordeaux_weather_data.csv", sep = ',' )
head(BORdf,3)
```

```{r}
ANTdf<-read.csv("C:/Users/Pham Antoine/Antibes_weather_data.csv", sep = ',')
head(ANTdf,3)
```

```{r}
library(funModeling)
df_status(BORdf)
```


```{r}
str(BORdf)
```

```{r}
summary(BORdf)
```

```{r}
df_status(ANTdf)
```

```{r}
ANTdf$Humidity<-as.integer(ANTdf$Humidity)
```

```{r}
str(ANTdf)
```

```{r}
summary(ANTdf)
```
Checking missing values & outliers
```{r}
subset(ANTdf,TemperatureC< -10)
```

```{r}
unique(BORdf$Conditions)
```
VISUALIZATION OF TEMPERATURE & HUMIDITY DISTRIBUTION IN BORDEAUX

```{r, fig.width = 12, fig.height = 8}
library(plyr)
tm <- ddply(BORdf, "Conditions", summarise, grp.mean=mean(TemperatureC))
ggplot(BORdf,aes(x=TemperatureC))+ 
  geom_density(aes(group=Conditions, colour=Conditions, fill=NULL), alpha=0.8)+
  geom_vline(data=tm, aes(xintercept=grp.mean, color=Conditions),linetype="dashed")

```

```{r,fig.width = 12, fig.height = 8}
tm1 <- ddply(BORdf, "Conditions", summarise, grp.mean=mean(Humidity))
ggplot(BORdf,aes(x=Humidity))+ 
  geom_density(aes(group=Conditions, colour=Conditions, fill=NULL), alpha=0.8)+
  geom_vline(data=tm1, aes(xintercept=grp.mean, color=Conditions),linetype="dashed")

```

 
VISUALIZATION OF TEMPERATURE & HUMIDITY DISTRIBUTION IN ANTIBES

```{r,fig.width = 12, fig.height = 8}
t=subset(ANTdf,TemperatureC> -10)# for ignoring records having outliers=-9999 as temperature
tn<- ddply(t, "Conditions", summarise, grp.mean=mean(TemperatureC))
ggplot(t,aes(x=TemperatureC))+ 
  geom_density(aes(group=Conditions, colour=Conditions, fill=NULL), alpha=0.8)+
  geom_vline(data=tn, aes(xintercept=grp.mean, color=Conditions),linetype="dashed")

```

```{r,fig.width = 12, fig.height = 8}
tn1<- ddply(ANTdf, "Conditions", summarise, grp.mean=mean(Humidity))
ggplot(ANTdf,aes(x=Humidity))+ 
  geom_density(aes(group=Conditions, colour=Conditions, fill=NULL), alpha=0.8)+
  geom_vline(data=tn1, aes(xintercept=grp.mean, color=Conditions),linetype="dashed")

```

```{r}
library(UsingR)
attach(BORdf)
plot(TemperatureC,Humidity,main=paste("Correlation: Temperature vs. Humidity in Bordeaux",signif(cor(TemperatureC,TemperatureC),2)))
```

```{r, fig.width = 12, fig.height = 8}
gg <- ggplot(BORdf, aes(x=Humidity, y=TemperatureC)) +
geom_point(aes(col=Conditions,size=0.001)) +
geom_smooth(method="loess", se=F) +
xlim(c(0, 110)) +
ylim(c(-10, 40)) +
labs(subtitle="Correlation :Temperature- Humidity-Conditions in Bordeaux",
y="Temperature °C",
x="Humidity",
title="Scatterplot",
caption = "Source: scraped from R package 'weatherData'_https://www.wunderground.com")
plot(gg)
```

```{r,fig.width = 8, fig.height = 6}
barplot(table(BORdf$Conditions), 
        main= 'Conditions',las=2,col = BORdf$Conditions,cex.axis=0.8,cex.names=1)#las=2 for ajusting the angle of texts:https://stackoverflow.com/questions/10286473/rotating-x-axis-labels-in-r-for-barplot
#, 
 #       col= c(len(BORdf$Conditions)))
#, legend.text =c("0 = Non","1 = Partielle","2 = Totale"),xlab="Type de resiliation", ylab="Nombre de contrats",ylim=c(0,150000), beside=TRUE,args.legend = list(x="topright",title= "Type", cex=0.6))
```

```{r}
table(BORdf$Conditions)
```

```{r}
library(UsingR)
attach(t)
plot(TemperatureC,Humidity,main=paste("Correlation: Temperature vs. Humidity in Antibes",signif(cor(TemperatureC,TemperatureC),2)))
```

```{r,fig.width = 12, fig.height = 8}
gq <- ggplot(t, aes(x=Humidity, y=TemperatureC)) +
geom_point(aes(col=Conditions,size=0.001)) +
geom_smooth(method="loess", se=F) +
xlim(c(0, 110)) +
ylim(c(-10, 40)) +
labs(subtitle="Correlation :Temperature- Humidity-Conditions in Antibes",
y="Temperature °C",
x="Humidity",
title="Scatterplot",
caption = "Source: scraped from R package 'weatherData'-https://www.wunderground.com")
plot(gq)
```

```{r}
barplot(table(ANTdf$Conditions), 
        main= 'Conditions',las=2,col = BORdf$Conditions,cex.axis=0.8,cex.names=1)
```


```{r}
table(ANTdf$Conditions)
```

```{r}
head(BORdf)
```

```{r}
as.POSIXct(BORdf$Time[1])
```

```{r}
library(lubridate)
hour(BORdf$Time[1])
```

```{r}
d=date(BORdf$Time[1])
d
```
```{r}
class(BORdf$Time)
```

Extracting under Date format from the date factor in the column 'Time' to a new column called "Date" 
```{r}
library(lubridate)
BORdf$Date<-as.Date(BORdf$Time)
tail(BORdf,2)
```
Extracting now the day of the week in a new column called ' DoW'

The function 'wday' returns the day of the week as a decimal number (01-07, Sunday is 1) or an ordered factor(Sunday is first).

```{r}
BORdf$DoW<-wday(BORdf$Date)
```

Extracting now the day of the month in a new column called ' DoM',teh week of the year to 'WoY', the month of the year to "MoY", the querter of the year to 'QoY'
```{r}
BORdf$DoM<-mday(BORdf$Date)
BORdf$WoY<-week(BORdf$Date)
BORdf$MoY<-month(BORdf$Date)
BORdf$QoY<-quarter(BORdf$Date)
tail(BORdf)
```

Creating a column called "schedule"" containing one of the following values: 'a_day_off', 'before_a_day_off', 'holidays' or 'no_event'
```{r}

BORdf$schedule<-ifelse(BORdf$Date=='2016-01-01'| BORdf$Date=='2016-03-28'|BORdf$Date =='2016-05-01'|BORdf$Date=='2016-05-08'|
                       BORdf$Date=='2016-05-16'| BORdf$Date=='2016-07-14'| BORdf$Date=='2016-08-15'|
                       BORdf$Date=='2016-11-01'|BORdf$Date=='2016-11-11'|BORdf$Date=='2016-12-25'|
                       BORdf$Date=='2017-01-01'| BORdf$Date=='2017-04-17'|BORdf$Date=='2017-05-01'| BORdf$Date=='2017-05-08'| BORdf$Date=='2017-05-25'|BORdf$Date=='2017-06-05'|BORdf$Date=='2017-07-14'| BORdf$Date=='2017-08-15'|BORdf$Date=='2017-11-01'|BORdf$Date=='2017-11-11'|BORdf$Date=='2017-12-25',
                       'a_day_off',
                       ifelse(BORdf$Date+1=='2016-01-01'| BORdf$Date+1=='2016-03-28'|BORdf$Date+1 =='2016-05-01'| BORdf$Date+1=='2016-05-08'|
                       BORdf$Date+1=='2016-05-16'| BORdf$Date+1=='2016-07-14'| BORdf$Date+1=='2016-08-15'|
                       BORdf$Date+1=='2016-11-01'|BORdf$Date+1=='2016-11-11'|
                       BORdf$Date+1=='2017-01-01'| BORdf$Date+1=='2017-05-01'| BORdf$Date+1=='2017-05-08'|
                       BORdf$Date+1=='2017-05-16'| BORdf$Date+1=='2017-07-14'| BORdf$Date+1=='2017-08-15'|BORdf$Date+1=='2017-11-01'|BORdf$Date+1=='2017-11-11',
                       'before_aday_off',
                       ifelse(BORdf$Date <'2016-01-04'| BORdf$Date >'2016-02-13'& BORdf$Date < '2016-02-29'| BORdf$Date >'2016-04-09'&
                       BORdf$Date <'2016-04-25'| BORdf$Date>'2016-05-04'& BORdf$Date<'2016-05-09'|
                       BORdf$Date>'2016-07-05'&BORdf$Date<'2016-09-01'|
                       BORdf$Date>'2016-10-19'& BORdf$Date<'2016-11-03'|BORdf$Date>'2016-12-17'&BORdf$Date<'2017-01-03'|
                       BORdf$Date>'2017-02-18'& BORdf$Date<'2017-03-06'| BORdf$Date>'2017-04-15'&BORdf$Date<'2017-05-02'|BORdf$Date>'2017-05-24'&BORdf$Date<'2017-05-29'|
                         BORdf$Date>'2017-07-08'&BORdf$Date<'2017-09-04',
                       'holidays',
                       'no_event')
                       ))
```

*Note: can use 'with' to avoid repeting Bordf$..

Bordf$schedule<-with(ifelse(Date==...,
                              a_day_off,
                              ifelse( Date+1,.....
                              'befor_a_day_off',...????????????????????????????????????????????????????????????????????????????????????????????????????????????????????????????????

```{r}
BORdf$schedule<-as.factor(BORdf$schedule)
head(BORdf,3)
```

```{r}
# check
subset(BORdf,Date=='2017-07-30')$schedule[1:5]

```
Working with date( some examples)
```{r}
l=substring(BORdf$DateUTC[1],12,19)
l= as.factor(l)
print(l)
```

```{r}
class(l)
```

```{r}
l=as.factor(l)
class(l)
```

```{r}
l<-as.POSIXct(l,format="%H:%M:%S")
l
# if date non given, the current date is showed by default
```

```{r}
class(l)
```


```{r}
hour(l)
```
LET'S KEEP BUIDING OUR " CONTEXT" DATA SET 
Adding a column named 'hour' indicating, the class of this feature will be integrer

```{r}
BORdf$H<- hour(BORdf$Time)
colnames(BORdf)

class(BORdf$H)
```
Now, create a column containing !h:m:s, the class of this feature will be factor
```{r}
BORdf$HMS<-substring(BORdf$Time,12,19)# take from the 12th to 19th elements of the variable 'Time'
class(BORdf$HMS)
```
converting this feature to factor   
```{r}
BORdf$HMS<-as.factor(BORdf$HMS)
head(BORdf,1)
```

```{r}
table(BORdf$HMS)
```

```{r}
summary(BORdf)
```

```{r}
colnames(BORdf)
```

```{r}
head(BORdf,2)

```

```{r,fig.width = 10,fig.height = 6}
ggplot(BORdf, aes(Humidity,TemperatureC))+ geom_point(aes(color = schedule)) +
scale_x_continuous("Humidity", breaks = seq(0,100,by = 5))+
scale_y_continuous("TemperatureC", breaks = seq(-10,110,by = 5))+
labs(subtitle="Correlation :Temperature vs. Humidity by the type of day in BORDEAUX")+
theme_bw() + labs(title="Scatterplot")

```

```{r,,fig.width = 10,fig.height = 6}
# this code is from : "7 Visualizations You Should Learn in R | Rbloggers"

ggplot(BORdf, aes(Humidity,TemperatureC)) + geom_point(aes(color = schedule)) +
scale_x_continuous("Humidity", breaks = seq(0,110,10))+
scale_y_continuous("TemperatureC", breaks = seq(-10,40,5))+
theme_bw() + labs(title="Scatterplot") + facet_wrap( ~ schedule)

```

```{r}
sub1=subset(BORdf,TemperatureC < 0,select = c('schedule','DoW','MoY'))
table(sub1$schedule,sub1$DoW)
```

```{r}
library(dplyr)
sub1<-tbl_df(sub1)# converting sub1 from df to table with 'dplyr' package, using filter to extract data ( no need subset anymore)
filter(sub1,schedule=='no_event'&DoW==7)
```
```{r}
write.csv(BORdf, file = 'Bordeaux_context.csv',row.names = F)
```






Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
